Seasonality of submesoscale dynamics in the Kuroshio Extension

### 2016-08-17 writeup/rocha_etal.tex
Adds new references

%
%
%\begin{figure*}[ht]
%  \begin{center}
%    \includegraphics[width=.65\textwidth]{figs/fig5.pdf}
% \caption{\small Surface (horizontal) kinetic energy (a) and sea-surface height variance wavenumber spectra (b)
% in the 1/48$^\circ$ simulation. Solid lines
% are spectra based on hourly snapshots, and dashed lines are spectra based on daily-averaged
% fields.}
% \label{fig5}
% \end{center}
%\end{figure*}
%
%
%Figure \ref{fig5}a depicts the horizontal wavenumber spectra of surface KE in
%April and October. At scales larger than 20 km, the April and October spectra based on hourly
%velocity snapshots (solid lines) are nearly indistinguishable from each other within 95$\%$ confidence level.
%Consistent with the results of
%\citet{rocha_etal2016}, who analyzed a 3-month output of the LLC4320 simulation in Drake Passage,
%there is significant high-frequency variabilty at submesoscales. Daily-averaging
%the velocity field suppresses spatial variability at scales smaller than about 250
%km, both in April and October (compare solid lines against dashed lines in Figure
%\ref{fig5}a). But this suppression is dramatic in October, when the inertia-gravity waves peak. At scales
%smaller than 100 km, 39$\%$ of the surface KE in April is accounted for by super-inertial
%flows as opposed to 79$\%$ in October. The seasonality of subinertial submesoscale flows
%is strong, consistent
%with the results of \citet{sasaki_etal2014}, which are
%based on daily-averaged velocity fields of a different model without tidal forcing
%(P. Klein, personal communication).
%
%The inertia-gravity waves project on the sea-surface.
%There is a dramatic difference between the spectra
%based on hourly and daily-averaged sea-surface height in October (Figure \ref{fig5}b): at scales smaller than 100 km, the spectra
%of hourly sea-surface height roughly follows a $-2$ power-law, whereas the spectra of daily-averaged
%sea-surface height roughly follows a $-5$ power-law. At these scales, 33$\%$ of the sea-surface height variance in April is accounted for by super-inertial
%flows as opposed to 83$\%$ in October. Curiously, the out-of-phase seasonal cycle of submesoscale
%turbulence and near-surface submesoscale inertia-gravity waves conspire to yield weak seasonality
%in the spectra of KE and sea-surface height variance based on hourly fields.
%
%
\section{Summary and Conclusion}
Our work adds to recent modeling \citep{sasaki_etal2014}
and observational \citep{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in
submesoscale turbulence. In particular, our main finding is that in global simulations with embedded tides the
near-surface submesoscale inertia-gravity waves in the Kuroshio Extension undergo
a strong seasonal cycle that is out of phase with the seasonal cycle of
submesoscale turbulence. We conjecture that the summertime dominance of inertia-gravity waves
 \citep{callies_etal2015} is a consequence both of suppression of submesoscale turbulence and
enhancement of inertia-gravity waves due to re-stratification of
the upper ocean.

\cite{dasaro1978} showed that the velocity of linear internal waves
in the mixed layer strongly depends on the density jump at the mixed layer
base, with largest mixed layer velocities when the jump is strongest.
In summer, the shallow mixed layer overlays a strong seasonal pycnocline,
and thus the internal waves projection onto the mixed layer may be stronger
according to \cite{dasaro1978}'s arguments. An
alternative explanation is that the shape of the baroclinic modes
changes seasonally. In particular the baroclinic modes are
significantly more surface intensified in late summer/early fall
 (supporting information).
If the internal wave source has weak seasonal dependence, then a strong
seasonal cycle in the near-surface expression of internal waves is expected.
Internal tide generation does not show a seasonal modulation \citep[e.g.,][]{alford2003},
and therefore one expects a strong seasonality in the near-surface
expression of internal tides and
other submesoscale inertia-gravity waves generated through internal tide interactions.
Near-inertial wave generation peaks in winter \citep[e.g.,][]{alford2003}, but
those waves project onto large horizontal scales \citep[e.g, ][]{qi_etal1995}.

@@

\bibliographystyle{agufull08}
%\bibliography{rocha_etal}

@@

%% ------------------------------------------------------------------------ %%
%%  REFERENCE LIST AND TEXT CITATIONS
%
% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
\begin{thebibliography}{18}
\providecommand{\natexlab}[1]{#1}
\expandafter\ifx\csname urlstyle\endcsname\relax
  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
  \urlstyle{rm}\Url}\fi

\bibitem[{\textit{Alford}(2003)}]{alford2003}
Alford, M.~H. (2003), {Redistribution of energy available for ocean mixing by
  long-range propagation of internal wave}, \textit{Nature},
  \textit{423}(6936), 159--162.

\bibitem[{\textit{Buckingham et~al.}(2016)\textit{Buckingham, Naveira~Garabato,
  Thompson, Brannigan, Lazar, Marshall, George~Nurser, Damerell, Heywood, and
  Belcher}}]{buckingham_etal2016}
Buckingham, C.~E., A.~C. Naveira~Garabato, A.~F. Thompson, L.~Brannigan,
  A.~Lazar, D.~P. Marshall, A.~George~Nurser, G.~Damerell, K.~J. Heywood, and
  S.~E. Belcher (2016), {Seasonality of submesoscale flows in the ocean surface
  boundary layer}, \textit{Geophysical Research Letters}, \textit{43}(5),
  2118--2126.

\bibitem[{\textit{B{\"u}hler et~al.}(2014)\textit{B{\"u}hler, Callies, and
  Ferrari}}]{buhler_etal2014}
B{\"u}hler, O., J.~Callies, and R.~Ferrari (2014), {Wave--vortex decomposition
  of one-dimensional ship-track data}, \textit{Journal of Fluid Mechanics},
  \textit{756}, 1007--1026.

\bibitem[{\textit{Callies et~al.}(2015)\textit{Callies, Ferrari, Klymak, and
  Gula}}]{callies_etal2015}
Callies, J., R.~Ferrari, J.~M. Klymak, and J.~Gula (2015), {Seasonality in
  submesoscale turbulence}, \textit{Nature communications}, \textit{6}.

\bibitem[{\textit{Callies et~al.}(2016)\textit{Callies, Flierl, Ferrari, and
  Fox-Kemper}}]{callies_etal2016}
Callies, J., G.~Flierl, R.~Ferrari, and B.~Fox-Kemper (2016), {The role of
  mixed-layer instabilities in submesoscale turbulence}, \textit{Journal of
  Fluid Mechanics}, \textit{788}, 5--41.

\bibitem[{\textit{Capet et~al.}(2008)\textit{Capet, McWilliams, Molemaker, and
  Shchepetkin}}]{capet_etal2008a}
Capet, X., J.~C. McWilliams, M.~J. Molemaker, and A.~Shchepetkin (2008),
  {Mesoscale to submesoscale transition in the California Current System. Part
  I: Flow structure, eddy flux, and observational tests}, \textit{Journal of
  Physical Oceanography}, \textit{38}(1), 29--43.

\bibitem[{\textit{D'Asaro}(1978)}]{dasaro1978}
D'Asaro, E.~A. (1978), {Mixed layer velocities induced by internal wave},
  \textit{Journal of Geophysical Research: Oceans}, \textit{83}(C5),
  2437--2438.

\bibitem[{\textit{Forget et~al.}(2015)\textit{Forget, Campin, Heimbach, Hill,
  Ponte, and Wunsch}}]{forget_etal2015}
Forget, G., J.-M. Campin, P.~Heimbach, C.~Hill, R.~Ponte, and C.~Wunsch (2015),
  {ECCO version 4: an integrated framework for non-linear inverse modeling and
  global ocean state estimation}, \textit{Geosci. Model Dev.}, \textit{8},
  3071--3104.

  \bibitem[{\textit{Marshall et~al.}(1997)\textit{Marshall, Hill, Perelman, and
    Adcroft}}]{marshall_etal1997}
  Marshall, J., C.~Hill, L.~Perelman, and A.~Adcroft (1997), Hydrostatic,
    quasi-hydrostatic, and nonhydrostatic ocean modeling, \textit{Journal of
    Geophysical Research: Oceans}, \textit{102}(C3), 5733--5752.

  \bibitem[{\textit{McWilliams}(2016)}]{mcwilliams2016}
  McWilliams, J.~C. (2016), {Submesoscale currents in the ocean}, in
    \textit{Proc. R. Soc. A}, vol. 472, p. 20160117, The Royal Society.

  \bibitem[{\textit{Menemenlis et~al.}(2008)\textit{Menemenlis, Campin, Heimbach,
    Hill, Lee, Nguyen, Schodlok, and Zhang}}]{menemenlis_etal2008}
  Menemenlis, D., J.-M. Campin, P.~Heimbach, C.~Hill, T.~Lee, A.~Nguyen,
    M.~Schodlok, and H.~Zhang (2008), ECCO2: High resolution global ocean and sea
    ice data synthesis, \textit{Mercator Ocean Quarterly Newsletter},
    \textit{31}, 13--21.

\bibitem[{\textit{Qi et~al.}(1995)\textit{Qi, De~Szoeke, Paulson, and
  Eriksen}}]{qi_etal1995}
Qi, H., R.~A. De~Szoeke, C.~A. Paulson, and C.~C. Eriksen (1995), {The
  structure of near-inertial waves during ocean storms}, \textit{Journal of
  Physical Oceanography}, \textit{25}, 2853--2871.

\bibitem[{\textit{Qiu et~al.}(2014)\textit{Qiu, Chen, Klein, Sasaki, and
  Sasai}}]{qiu_etal2014}
Qiu, B., S.~Chen, P.~Klein, H.~Sasaki, and Y.~Sasai (2014), {Seasonal mesoscale
  and submesoscale eddy variability along the North Pacific Subtropical
  Countercurrent}, \textit{Journal of Physical Oceanography}, \textit{44}(12),
  3079--3098.

\bibitem[{\textit{Richman et~al.}(2012)\textit{Richman, Arbic, Shriver,
  Metzger, and Wallcraft}}]{richman_etal2012}
Richman, J.~G., B.~K. Arbic, J.~F. Shriver, E.~J. Metzger, and A.~J. Wallcraft
  (2012), {Inferring dynamics from the wavenumber spectra of an eddying global
  ocean model with embedded tides}, \textit{Journal of Geophysical Research:
  Oceans}, \textit{117}(C12).

\bibitem[{\textit{Rocha et~al.}(2016)\textit{Rocha, Chereskin, Gille, and
  Menemenlis}}]{rocha_etal2016}
Rocha, C.~B., T.~K. Chereskin, S.~T. Gille, and D.~Menemenlis (2016),
  {Mesoscale to submesoscale wavenumber spectra in Drake Passage},
  \textit{Journal of Physical Oceanography}, \textit{46}, 601--620.

\bibitem[{\textit{Sasaki et~al.}(2014)\textit{Sasaki, Klein, Qiu, and
  Sasai}}]{sasaki_etal2014}
Sasaki, H., P.~Klein, B.~Qiu, and Y.~Sasai (2014), {Impact of oceanic-scale
  interactions on the seasonal modulation of ocean dynamics by the atmosphere},
  \textit{Nature communications}, \textit{5}.

\bibitem[{\textit{Shcherbina et~al.}(2013)\textit{Shcherbina, D'Asaro, Lee,
  Klymak, Molemaker, and McWilliams}}]{shcherbina_etal2013}
Shcherbina, A.~Y., E.~A. D'Asaro, C.~M. Lee, J.~M. Klymak, M.~J. Molemaker, and
  J.~C. McWilliams (2013), {Statistics of vertical vorticity, divergence, and
  strain in a developed submesoscale turbulence field}, \textit{Geophysical
  Research Letters}, \textit{40}(17), 4706--4711.

\bibitem[{\textit{Thompson et~al.}(2016)\textit{Thompson, Lazar, Buckingham,
  Naveira~Garabato, Damerell, and Heywood}}]{thompson_etal2016}
Thompson, A.~F., A.~Lazar, C.~Buckingham, A.~C. Naveira~Garabato, G.~M.
  Damerell, and K.~J. Heywood (2016), {Open-ocean submesoscale motions: A full
  seasonal cycle of mixed layer instabilities from gliders}, \textit{Journal of
  Physical Oceanography}, \textit{46}(4), 1285--1307.

\end{thebibliography}

% Or,
%
% If you use BiBTeX for your references, please use the agufull08.bst file (available at % ftp://ftp.agu.org/journals/latex/journals/Manuscript-Preparation/) to produce your .bbl
% file and copy the contents into your paper here.
%
% Follow these steps:
% 1. Run LaTeX on your LaTeX file.
%
% 2. Make sure the bibliography style appears as \bibliographystyle{agufull08}. Run BiBTeX on your LaTeX
% file.
%
% 3. Open the new .bbl file containing the reference list and
%   copy all the contents into your LaTeX file here.
%
% 4. Comment out the old \bibliographystyle and \bibliography commands.
%
% 5. Run LaTeX on your new file before submitting.
%
% AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.

>>>>

%
%
%\begin{figure*}[ht]
%  \begin{center}
%    \includegraphics[width=.65\textwidth]{figs/fig5.pdf}
% \caption{\small Surface (horizontal) kinetic energy (a) and sea-surface height variance wavenumber spectra (b)
% in the 1/48$^\circ$ simulation. Solid lines
% are spectra based on hourly snapshots, and dashed lines are spectra based on daily-averaged
% fields.}
% \label{fig5}
% \end{center}
%\end{figure*}
%
%
%Figure \ref{fig5}a depicts the horizontal wavenumber spectra of surface KE in
%April and October. At scales larger than 20 km, the April and October spectra based on hourly
%velocity snapshots (solid lines) are nearly indistinguishable from each other within 95$\%$ confidence level.
%Consistent with the results of
%\citet{rocha_etal2016}, who analyzed a 3-month output of the LLC4320 simulation in Drake Passage,
%there is significant high-frequency variabilty at submesoscales. Daily-averaging
%the velocity field suppresses spatial variability at scales smaller than about 250
%km, both in April and October (compare solid lines against dashed lines in Figure
%\ref{fig5}a). But this suppression is dramatic in October, when the inertia-gravity waves peak. At scales
%smaller than 100 km, 39$\%$ of the surface KE in April is accounted for by super-inertial
%flows as opposed to 79$\%$ in October. The seasonality of subinertial submesoscale flows
%is strong, consistent
%with the results of \citet{sasaki_etal2014}, which are
%based on daily-averaged velocity fields of a different model without tidal forcing
%(P. Klein, personal communication).
%
%The inertia-gravity waves project on the sea-surface.
%There is a dramatic difference between the spectra
%based on hourly and daily-averaged sea-surface height in October (Figure \ref{fig5}b): at scales smaller than 100 km, the spectra
%of hourly sea-surface height roughly follows a $-2$ power-law, whereas the spectra of daily-averaged
%sea-surface height roughly follows a $-5$ power-law. At these scales, 33$\%$ of the sea-surface height variance in April is accounted for by super-inertial
%flows as opposed to 83$\%$ in October. Curiously, the out-of-phase seasonal cycle of submesoscale
%turbulence and near-surface submesoscale inertia-gravity waves conspire to yield weak seasonality
%in the spectra of KE and sea-surface height variance based on hourly fields.
%
%
\section{Discussion and conclusion}
Our work adds to recent modeling \citep{sasaki_etal2014}
and observational \citep{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in
submesoscale turbulence. In particular, our main finding is that in global simulations with embedded tides the
near-surface submesoscale inertia-gravity waves in the Kuroshio Extension undergo
a strong seasonal cycle that is out of phase with the seasonal cycle of
submesoscale turbulence. We conjecture that the summertime dominance of inertia-gravity waves
 \citep{callies_etal2015} is a consequence both of suppression of submesoscale turbulence and
enhancement of inertia-gravity waves due to re-stratification of
the upper ocean.

\cite{dasaro1978} showed that the velocity of linear internal waves
in the mixed layer strongly depends on the density jump at the mixed layer
base, with largest mixed layer velocities when the jump is strongest.
In summer, the shallow mixed layer overlays a strong seasonal pycnocline,
and thus the internal waves projection onto the mixed layer may be stronger
according to \cite{dasaro1978}'s arguments. An
alternative explanation is that the shape of the baroclinic modes
changes seasonally (supporting information).

If the internal wave source has weak seasonal dependence, then a strong
seasonal cycle in the near-surface expression of internal waves is expected.
Internal tide generation does not show a seasonal modulation \citep[e.g.,][]{alford2003},
and therefore one expects a strong seasonality in the near-surface
expression of internal tides and
other submesoscale inertia-gravity waves generated through internal tide interactions.
Windy near-inertial wave (NIW) generation is bursty and peaks in winter \citep[e.g.,][]{alford2003}.
While NIWs are generated at large scales, mesoscale turbulence can
refract and disperse those waves in few inertial periods, thereby generating
submesoscale (10-100 km) near-inertial waves \citep{young_benjelloul1997}. But in our
simulations the wavy seasonality in submesoscale surface KE and SSH variance is mostly accounted for
by inertia-gravity waves with periods shorter than 12.4 h (Figure \ref{fig4}a-b).

@@

\bibliographystyle{agufull08}
\bibliography{rocha_etal}

@@

%% ------------------------------------------------------------------------ %%
%%  REFERENCE LIST AND TEXT CITATIONS
%
% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% Or,
%
% If you use BiBTeX for your references, please use the agufull08.bst file (available at % ftp://ftp.agu.org/journals/latex/journals/Manuscript-Preparation/) to produce your .bbl
% file and copy the contents into your paper here.
%
% Follow these steps:
% 1. Run LaTeX on your LaTeX file.
%
% 2. Make sure the bibliography style appears as \bibliographystyle{agufull08}. Run BiBTeX on your LaTeX
% file.
%
% 3. Open the new .bbl file containing the reference list and
%   copy all the contents into your LaTeX file here.
%
% 4. Comment out the old \bibliographystyle and \bibliography commands.
%
% 5. Run LaTeX on your new file before submitting.
%
% AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.
